Analysis for Revision 2

### 2024-02-21 R/Analysis_v1.2_mash_using_greedy_covar.qmd
feat: QTL analysis using phenotypes from new covariance matrices, incorporated into Figure 3

---
title: "Analysis v1.2 mash"
author: "Alice MacQueen"
date: "2023-08-01"
output: html_document
---

@@

# Installation reminders for any reinstalls
On debian: `apt install libgsl-dev` to install the GNU Scientific Library, zB.

For mashr: Sometimes install.packages("RcppGSL") is also needed.
#

Sam used a greedy mash algorithm to find the best set of covariance matrices to include, using 19K randomly associated variants.

Now use the best set of covariance matrices on 19K strongly associated variants.

Step 0. Set up.
  0a. Look at phenotype distribution & heritability; run GWAS to get univariate SNP effect estimates & standard errors. Already done.
  0b. Make hypothesis covariance matrices based on weather data from 2018 - 2020. For cumulative weather variables, use a range of days, 1 - 20, say. Come back and add more if 20 is chosen by greedy mash.
Step 1. Run greedy mash to find best set of covariance matrices that explain SNP effect variation across environments, using 19K 'randomly associated variants'. Go back to 0b. if high number of day covar matrices tend to be chosen by greedy mash.
Step 2. Use covariances from (1) to run mash on 19K 'strongly associated variants'. 
3. Look at loadings on covariance matrices and meanings of covariance matrices as the writeup (Figure 2, 3; Table 1). Look at amount of antagonistic pleiotropy vs same sign significant effects.
4. Repeat the QTL analysis... with Li.

How much of a lift is it to re-run the greedy mash algorithm? If we, or if reviewers are thinking about this in a Jiaming Yu sense, I think the next logical question to ask is 'Why these covariance matrices'? The overly honest answer for the current set is that there were computational limits to running mash with so many covariance matrices, so I chose a few arbitrary time points for each weather variable we thought might be a environmental driver. But now we have introduced a method to choose between highly correlated covariance matrices, perhaps we could do better. 
2.

# ----
# Generate weather variables

# Covar 1-28d weather
As of 2023-09-11, want to remake these per reviewer revision to have h2 on the diagonal instead of the CV. Need to estimate h2 for these new covar matrices, probably using rrBLUP.

## load phe & wea data
```{r}
phe_ten_sites <- readRDS(here("data", "Phenotypes_2018_to_2021_10_sites.rds"))
sites <- readRDS("~/Github/pvdiv-phenology-gxe/data/sites.rds")
jday_phe <- readRDS("~/Github/pvdiv-phenology-gxe/data/Julian_date_GDD_and_rainfall_phenotypes_4wcr_and_gwas.rds")
wea_ten_sites <- readRDS(file = here("data", "Daily_weather_2018_to_2021_10_sites.rds"))

@@

@@

```{r}
wea_ten_sites
```

# Define phenotypes
```{r}
Tbase <- 12

@@

## save weather-based phenotypes
```{r}
gr_ten_sites %>%
  saveRDS(file = here("data", "Weather_related_phe_GR50_2019.rds"))
fl_ten_sites %>%
  saveRDS(file = here("data", "Weather_related_phe_FL50_2019.rds"))
```
# --- can start here
# Covar csv for greedy mash algorithm

@@

phe_hyp <- c("cgdd_12c_10d", "cgdd_12c_18d", "cgdd_12c_5d", 
             "tave_5d", "tave_10d", "tave_18d")
phe_hyp <- c("FL50", "GR50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl", 
             "crain_gr2fl", "crain_1d", "crain_3d", "crain_5d")

### load weather phenotypes
```{r}
gr_ten_sites <- readRDS(file = here("data", 
                                    "Weather_related_phe_GR50_2019.rds"))
fl_ten_sites <- readRDS(file = here("data", 
                                    "Weather_related_phe_FL50_2019.rds"))
```

@@

## Calculate heritability for diagonal
I don't have access to ASREML any more, so need a different package to do this.

@@

Use the kinship matrix I've already calculated (using the van Raden method, using switchgrassGWAS::pvdiv_kinship()) using the SNPs as the K matrix.
```{r}
library(rrBLUP)
k_full <- read_rds("~/Github/pvdiv-genome/tensite_twoyear/Kinship_van_Raden_630_individuals_SNPs_r2_20percent.rds")
fl_ten_sites
colnames(k_full)

@@

### Loop to calculate covar for flowering
```{r}
metadata <- read_rds(here("data", "metadata.rds"))
subpop_v2 <- list(all = c("Gulf", "Midwest"), midwest = "Midwest",
                  gulf = "Gulf")
year_v = c(2019, 2020, 2021)

@@

```
Another error: "Error in mash(data, covmats) : 
  All U_scaled matrices should be positive semi-definite

### Loop to calculate covar for greenup
```{r}
gr_ten_sites <- gr_ten_sites %>%
  left_join(select(metadata, PLANT_ID, SUBPOP)) %>%
  rename(Year = YEAR)

@@

#### flowering: 6 sites
MI, MO, SD, TX1, TX2, TX3
```{r flowering midwest}

@@

#### greenup: 7 sites
MI, MO, NE, OK, TX1, TX2, TX3
```{r}
phe_hyp <- colnames(gr_ten_sites)[8:50]
i=1
k=2

@@

```{r}
workingdir <- here("analysis", "mash_greedy_algorithm", 
                                   "mash_h2_diag_covar", "midwest_greenup")
pheno = "greenup"
betas = read.table(here(workingdir, "combined_mash_effects", paste0( pheno,'.midwest.betas.random.txt')))
stderrs = read.table(here(workingdir, "combined_mash_stderrs", paste0( pheno,'.midwest.stderrs.random.txt')))

@@

#### Visualize correlation matrices
```{r}

i=7
switchgrassGWAS::mash_plot_pairwise_sharing(corrmatrix = U_hyp_2019[[i]],
                                            reorder = FALSE)
switchgrassGWAS::mash_plot_pairwise_sharing(corrmatrix = U_hyp_2019[[i+1]],
                                            reorder = FALSE)
switchgrassGWAS::mash_plot_pairwise_sharing(corrmatrix = U_hyp_2019[[i+2]],
                                            reorder = FALSE) 
names(U_hyp_2019)[[i+2]]
```

# ---------------
# Info to run greedy mash algo

@@

 '.singleton.*'
'.identity.*'
'.simple_het_*.*'
 '.equal_effects.*'
'.*PCA*'

@@

When I change to my inputs for effects & stderrs, I will need to adjust midwest covar & effects appropriately to reflect only 6 (flowering) and 7 (greenup) rows & columns used; the other GWAS were dropped as too low power or due to substantial phenotype data loss, & I think including them would be misleading, likely.

### Next steps
Redo Midwest: Remake Midwest covar matrices, including hypothesis, singleton, identity, simple het, equal effects, and data-driven matrices. Make mash_greedy.R variants for changed input dir & for different Midwest column numbers. 
Run two new run.mash.greedy.search.py commands:
python run.mash.greedy.search.py -p greenup -c midwest -e random
python run.mash.greedy.search.py -p flowering -c midwest -e random

@@

I only really need the random ones here, run mash again on the strong effects below.
```{r}
popkey = list(all = "Gulf_and_Midwest", midwest = "Midwest", gulf = "Gulf")
phekey = list(greenup = "GR50", flowering = "FL50")
# i=2; j=1

@@

```{r}
for (i in 1:35) {
  if (!dir.exists(here("analysis", "mash_greedy_algorithm", "mash_h2_diag_covar", "midwest_inputs", "likelihoods", paste0("step", i)))) {
    dir.create(here("analysis", "mash_greedy_algorithm", "mash_h2_diag_covar", "midwest_inputs", "likelihoods", paste0("step", i)))
  }
}
```

@@

# -----------------------------
# Run mash

```{r}
library(tidyverse)
library(bigsnpr)
library(switchgrassGWAS)
library(mashr)
library(rlist)
library(here)

@@

# Remake mash figures

## Load mash objects
```{r}
source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
mashdir <- here("analysis", "mash_greedy_algorithm", "mash_h2_diag_covar")

@@

## Figure 2
```{r}
get_cov_type <- function(m, summarise = FALSE, subpop = FALSE){
  cov_df <- mash_plot_covar(m)
  cov_type <- cov_df$covar_df %>% 
    mutate(Type = case_when(grepl("PCA", `Covariance Matrix`) ~ "data-driven",
                            grepl("singleton", `Covariance Matrix`) ~
                              "canonical",
                            grepl("simple_het", `Covariance Matrix`) ~
                              "canonical",
                            grepl("no_effects", `Covariance Matrix`) ~
                              "canonical",
                            grepl("identity", `Covariance Matrix`) ~
                              "canonical",
                            grepl("equal_effects", `Covariance Matrix`) ~
                              "canonical",
                            grepl("U.mat$", `Covariance Matrix`) ~
                              "hypothesized",
                            TRUE ~ "other"
                            )) 
  if (summarise == TRUE) {
    cov_type <- cov_type %>%
      group_by(Type) %>%
      summarise(Mass = sum(Mass))
  }
  return(cov_type)
}

get_cov_type(m_gulf_gr_hyp, summarise = TRUE, subpop = TRUE)
get_cov_type(m_gulf_fl_hyp, summarise = TRUE, subpop = TRUE)
get_cov_type(m_midw_gr_hyp, summarise = FALSE, subpop = TRUE)
get_cov_type(m_midw_fl_hyp, summarise = TRUE, subpop = TRUE)
get_cov_type(m_gam_gr_hyp, summarise = TRUE, subpop = TRUE)
get_cov_type(m_gam_fl_hyp, summarise = TRUE, subpop = TRUE)
```

@@

cov_gg$ggobject + 
  ggtitle("Gulf greenup", subtitle = "64.5% SNP mass")
cov_gf$ggobject + 
  ggtitle("Gulf flowering", subtitle = "33.0% SNP mass")
cov_mg$ggobject + 
  ggtitle("Midwest greenup", subtitle = "0% SNP mass")
cov_mf$ggobject + 
  ggtitle("Midwest flowering", subtitle = "22.6% SNP mass")
cov_gamg$ggobject + 
  ggtitle("Gulf and Midwest greenup", subtitle = "32.7% SNP mass")
cov_gamf$ggobject + 
  ggtitle("Gulf and Midwest flowering", subtitle = "0% SNP mass")
```
### covar

@@

```{r build panels of barplots for fig 2}
cov_gr1 <- mash_plot_Ulist(m_gulf_gr_hyp, range = 2, limits = TRUE)
legend <- get_legend(cov_gr1) 
fig_eqeff <- cov_gr1$greenup.simple_het_2_df %>%
  get_covar_fig2() +
  labs(subtitle = "simple heterozygousity")

@@

```{r build figure 2}
ggdraw() +
  draw_plot(pa01, x = -0.02, y = 0.4, width = 0.1, height = 0.18) +
  draw_plot(pa02, x = -0.02, y = 0.6, width = 0.1, height = 0.18) +
  #draw_plot(pa03, x = -0.02, y = 0.6, width = 0.1, height = 0.18) +
  #draw_plot(pa04, x = 0.1, y = 0.6, width = 0.1, height = 0.18) +
  draw_plot(pa05, x = -0.02, y = 0.8, width = 0.1, height = 0.18) +
  draw_plot(pa06, x = -0.02, y = 0, width = 0.1, height = 0.18) +
  draw_plot(pa07, x = -0.02, y = 0.2, width = 0.1, height = 0.18) +
  #draw_plot(pa08, x = 0.2, y = 0, width = 0.1, height = 0.18) +
  draw_plot(pa09, x = 0.13, y = 0.6, width = 0.1, height = 0.18) +
  draw_plot(pa10, x = 0.13, y = 0.2, width = 0.1, height = 0.18) +
  draw_plot(pa11, x = 0.13, y = 0.4, width = 0.1, height = 0.18) +
  draw_plot(pa12, x = 0.13, y = 0, width = 0.1, height = 0.18) +
  draw_plot(pa13, x = 0.13, y = 0.8, width = 0.1, height = 0.18) +
  draw_plot(f2a, x = 0.3, y = 0.47, width = 0.55, height = 0.53) +
  #draw_plot(legend, x = 0.15, y = 0.34, width = 0.1, height = 0.2) +
  draw_plot(f2b, x = 0.85, y = 0.47, width = 0.15, height = 0.53) +
  draw_plot(f2c, x = 0.3, y = 0, width = 0.55, height = 0.47) +
  draw_plot(f2d, x = 0.85, y = 0, width = 0.15, height = 0.47)

save_plot(filename = "../manuscript/Coauthors/PNAS/Revision/Figure_2_Covariance_matrices_and_posterior_weights.svg", plot = last_plot(), base_height = 6, base_width = 8.5, )
```

@@

Change these to look at S_AP, S_DS, S_2_no; S_1_row and S_1_col separately
```{r}

@@

get_gxe_summary <- function(gxe_list, n_cond = 8, n_sep = 2) {
  AP <- one_gxe_summary(gxe_list$S_AP, n_cond = n_cond, n_sep = n_sep) %>%
    rename(Sign = n)
  DS <- one_gxe_summary(gxe_list$S_DS, n_cond = n_cond, n_sep = n_sep) %>%
    rename(Magnitude = n)
  ND <- one_gxe_summary(gxe_list$S_2_no, n_cond = n_cond, n_sep = n_sep) %>%
    rename(None = n)
  gxe_n <- AP %>%
    full_join(DS, by = "Region_comparison") %>%
    full_join(ND, by = "Region_comparison") %>%
    pivot_longer(cols = -Region_comparison, names_to = "Type_of_GxE", 
                 values_to = "n") %>%
    pivot_wider(names_from = "Region_comparison", values_from = "n")
  gxe_frac <- gxe_n %>%
    mutate(North_North = round(North_North/sum(North_North), digits = 4),
           North_Texas = round(North_Texas/sum(North_Texas), digits = 4),
           Texas_Texas = round(Texas_Texas/sum(Texas_Texas), digits = 4))
  return(list(gxe_counts = gxe_n, gxe_fractions = gxe_frac))
}

@@

```{r}
get_gxe_summary(gxe_gg) 
get_gxe_summary(gxe_gf) 
get_gxe_summary(gxe_mg, n_cond = 7) 
get_gxe_summary(gxe_mf, n_cond = 6) 
get_gxe_summary(gxe_gamg, n_sep = 4) 
get_gxe_summary(gxe_gamf, n_sep = 4) 
```

@@

```{r}
mash_plot_Ulist(m_midw_fl_hyp)
mash_plot_manhattan_by_condition(m_midw_fl_hyp)
```

@@

@@

>>>>

---
title: "Analysis for Revision 2"
author: "Alice MacQueen"
format: 
  html:
    code-fold: true
    code-annotations: hover
    execute: 
      echo: false
      cache: true
editor: visual
toc: true
number-sections: true
---

::: callout-important
## Analysis Version 1.2

As of 2024-02-07, this file was a copy of the Rmd file of the same name. After 2024-02-07, this Quarto document will be modified and kept up to date for this analysis version.
:::

@@

::: callout-tip
## Install reminders for reinstalls

On debian: `apt install libgsl-dev` to install the GNU Scientific Library, zB.

For mashr: Sometimes install.packages("RcppGSL") is also needed.
:::

## Overview

Sam developed a greedy mash algorithm to find the best set of covariance matrices to include. Alice used this algorithm on 19K randomly associated variants using a large set of covariance matrices with h\^2 on their diagonals.

Then, Alice used the best set of covariance matrices on 19K strongly associated variants.

Outline of the current analysis:

1.  Set up. Look at phenotype distribution & heritability; run GWAS to get univariate SNP effect estimates & standard errors. Done elsewhere.

2.  Make hypothesis covariance matrices based on weather data from 2018 - 2020. For cumulative weather variables, use a range of days: every day for the first week, and every week until 28 days. Come back and add more if 28 is chosen by greedy mash.

3.  Run greedy mash to find best set of covariance matrices that explain SNP effect variation across environments, using 19K 'randomly associated variants'. Go back to (1) if high number of day covar matrices tend to be chosen by greedy mash.

4.  Use covariances from (3) to run mash on 19K 'strongly associated variants'.

5.  Look at loadings on covariance matrices and meanings of covariance matrices as the writeup (Figure 2, 3; Table 1). Look at amount of antagonistic pleiotropy vs same sign significant effects.

6.  Have Li repeat the QTL analysis and incorporate her results.

------------------------------------------------------------------------

## Step 1: GWAS

## Step 2: Covariance matrices

### Generate weather variables

### Covar 1-28d weather

As of 2023-09-11, want to remake these per reviewer revision to have h\^2 on the diagonal instead of the CV. Estimate h\^2 for these new covar matrices using rrBLUP.

### load phe & wea data

```{r}
#| eval: false

phe_ten_sites <- readRDS(here("data", "Phenotypes_2018_to_2021_10_sites.rds"))
sites <- readRDS("~/Github/pvdiv-phenology-gxe/data/sites.rds")
jday_phe <- readRDS("~/Github/pvdiv-phenology-gxe/data/Julian_date_GDD_and_rainfall_phenotypes_4wcr_and_gwas.rds")
wea_ten_sites <- readRDS(file = here("data", "Daily_weather_2018_to_2021_10_sites.rds"))

@@

#| eval: false
#| 

@@

```{r}
#| eval: false
wea_ten_sites
```

# Define phenotypes

```{r}
#| eval: false
#| 
Tbase <- 12

@@

## save weather-based phenotypes

```{r}
#| eval: false
gr_ten_sites %>%
  saveRDS(file = here("data", "Weather_related_phe_GR50_2019.rds"))
fl_ten_sites %>%
  saveRDS(file = here("data", "Weather_related_phe_FL50_2019.rds"))
```

# --- can start here

# Covar csv for greedy mash algorithm

@@

phe_hyp \<- c("cgdd_12c_10d", "cgdd_12c_18d", "cgdd_12c_5d", "tave_5d", "tave_10d", "tave_18d") phe_hyp \<- c("FL50", "GR50", "dyln_fl50", "dyln_change_sec", "cgdd_12c_gr2fl", "crain_gr2fl", "crain_1d", "crain_3d", "crain_5d")

### load weather phenotypes

```{r}
gr_ten_sites <- readRDS(file = here("data", 
                                    "Weather_related_phe_GR50_2019.rds"))
fl_ten_sites <- readRDS(file = here("data", 
                                    "Weather_related_phe_FL50_2019.rds"))
```

@@

## Calculate heritability for diagonal

I don't have access to ASREML any more, so need a different package to do this.

@@

Use the kinship matrix I've already calculated (using the van Raden method, using switchgrassGWAS::pvdiv_kinship()) using the SNPs as the K matrix.

```{r}
#| eval: false

library(rrBLUP)
k_full <- read_rds("~/Github/pvdiv-genome/tensite_twoyear/Kinship_van_Raden_630_individuals_SNPs_r2_20percent.rds")
fl_ten_sites
colnames(k_full)

@@

### Loop to calculate covar for flowering

```{r}
#| eval: false

metadata <- read_rds(here("data", "metadata.rds"))
subpop_v2 <- list(all = c("Gulf", "Midwest"), midwest = "Midwest",
                  gulf = "Gulf")
year_v = c(2019, 2020, 2021)

@@

### Loop to calculate covar for greenup

```{r}
#| eval: false
#| 
gr_ten_sites <- gr_ten_sites %>%
  left_join(select(metadata, PLANT_ID, SUBPOP)) %>%
  rename(Year = YEAR)

@@

#### flowering: 6 sites

MI, MO, SD, TX1, TX2, TX3

```{r flowering midwest}
#| eval: false

@@

#### greenup: 7 sites

MI, MO, NE, OK, TX1, TX2, TX3

```{r}
#| eval: false

phe_hyp <- colnames(gr_ten_sites)[8:50]
i=1
k=2

@@

```{r}
#| eval: false
workingdir <- here("analysis", "mash_greedy_algorithm", 
                                   "mash_h2_diag_covar", "midwest_greenup")
pheno = "greenup"
betas = read.table(here(workingdir, "combined_mash_effects", paste0( pheno,'.midwest.betas.random.txt')))
stderrs = read.table(here(workingdir, "combined_mash_stderrs", paste0( pheno,'.midwest.stderrs.random.txt')))

@@

#### Visualize correlation matrices

```{r}
#| eval: false
i=7
switchgrassGWAS::mash_plot_pairwise_sharing(corrmatrix = U_hyp_2019[[i]],
                                            reorder = FALSE)
switchgrassGWAS::mash_plot_pairwise_sharing(corrmatrix = U_hyp_2019[[i+1]],
                                            reorder = FALSE)
switchgrassGWAS::mash_plot_pairwise_sharing(corrmatrix = U_hyp_2019[[i+2]],
                                            reorder = FALSE) 
names(U_hyp_2019)[[i+2]]
```

# ---------------

## Step 3: Greedy Mash

# Info to run greedy mash algo

@@

'.singleton.*' '.identity.*' '.simple_het\_*.*' '.equal_effects.*' '.*PCA\*'

@@

When I change to my inputs for effects & stderrs, I will need to adjust midwest covar & effects appropriately to reflect only 6 (flowering) and 7 (greenup) rows & columns used; the other GWAS were dropped as too low power or due to substantial phenotype data loss, & including them would give misleading results.

### Next steps

Redo Midwest: Remake Midwest covar matrices, including hypothesis, singleton, identity, simple het, equal effects, and data-driven matrices. Make mash_greedy.R variants for changed input dir & for different Midwest column numbers. Run two new run.mash.greedy.search.py commands: python run.mash.greedy.search.py -p greenup -c midwest -e random python run.mash.greedy.search.py -p flowering -c midwest -e random

@@

I only really need the random ones here, run mash again on the strong effects below.

```{r}
#| eval: false
popkey = list(all = "Gulf_and_Midwest", midwest = "Midwest", gulf = "Gulf")
phekey = list(greenup = "GR50", flowering = "FL50")
# i=2; j=1

@@

```{r}
#| eval: false
for (i in 1:35) {
  if (!dir.exists(here("analysis", "mash_greedy_algorithm", "mash_h2_diag_covar", "midwest_inputs", "likelihoods", paste0("step", i)))) {
    dir.create(here("analysis", "mash_greedy_algorithm", "mash_h2_diag_covar", "midwest_inputs", "likelihoods", paste0("step", i)))
  }
}
```

@@

# -----------------------------

# Step 4: Run mash

```{r}
#| eval: false
#| 
#|

library(tidyverse)
library(bigsnpr)
library(switchgrassGWAS)
library(mashr)
library(rlist)
library(here)

@@

# Step 5: Redo writeup & figures

## Load mash objects

```{r}
source("~/Github/Functions_ggplot-theme-adjustments_2018-01-03.R")
mashdir <- here("analysis", "mash_greedy_algorithm", "mash_h2_diag_covar")

@@

## Figure 2

```{r}
get_cov_type <- function(m, summarise = FALSE, subpop = FALSE){
  cov_df <- mash_plot_covar(m)
  cov_type <- cov_df$covar_df %>% 
    mutate(Type = case_when(grepl("PCA", `Covariance Matrix`) ~ "data-driven",
                            grepl("singleton", `Covariance Matrix`) ~
                              "canonical",
                            grepl("simple_het", `Covariance Matrix`) ~
                              "canonical",
                            grepl("no_effects", `Covariance Matrix`) ~
                              "canonical",
                            grepl("identity", `Covariance Matrix`) ~
                              "canonical",
                            grepl("equal_effects", `Covariance Matrix`) ~
                              "canonical",
                            grepl("U.mat$", `Covariance Matrix`) ~
                              "hypothesized",
                            TRUE ~ "other"
                            )) 
  if (summarise == TRUE) {
    cov_type <- cov_type %>%
      group_by(Type) %>%
      summarise(Mass = sum(Mass))
  }
  return(cov_type)
}

get_cov_type(m_gulf_gr_hyp, summarise = TRUE, subpop = TRUE)
get_cov_type(m_gulf_fl_hyp, summarise = TRUE, subpop = TRUE)
get_cov_type(m_midw_gr_hyp, summarise = FALSE, subpop = TRUE)
get_cov_type(m_midw_fl_hyp, summarise = TRUE, subpop = TRUE)
get_cov_type(m_gam_gr_hyp, summarise = FALSE, subpop = TRUE)
get_cov_type(m_gam_fl_hyp, summarise = TRUE, subpop = TRUE)
```

@@

cov_gg$ggobject + 
  ggtitle("Gulf greenup", subtitle = "64.5% SNP mass")
cov_gf$ggobject + 
  ggtitle("Gulf flowering", subtitle = "33.0% SNP mass")
cov_mg$ggobject + 
  ggtitle("Midwest greenup", subtitle = "0% SNP mass")
cov_mf$ggobject + 
  ggtitle("Midwest flowering", subtitle = "22.6% SNP mass")
cov_gamg$ggobject + 
  ggtitle("Gulf and Midwest greenup", subtitle = "32.7% SNP mass")
cov_gamf$ggobject + 
  ggtitle("Gulf and Midwest flowering", subtitle = "0% SNP mass")
```

### covar

@@

```{r build panels of barplots for fig 2}
cov_gr1 <- mash_plot_Ulist(m_gulf_gr_hyp, range = 2, limits = TRUE)
legend <- get_legend(cov_gr1) 
fig_eqeff <- plot_covar_Ulist_fig2(Ulist_mg[[2]], mg = TRUE) +
  labs(subtitle = "simple heterozygousity")

@@

```{r build figure 2}

ggdraw() +
  draw_plot(pa01, x = -0.02, y = 0.4, width = 0.1, height = 0.18) +
  draw_plot(pa02, x = -0.02, y = 0.6, width = 0.1, height = 0.18) +
  #draw_plot(pa03, x = -0.02, y = 0.6, width = 0.1, height = 0.18) +
  #draw_plot(pa04, x = 0.1, y = 0.6, width = 0.1, height = 0.18) +
  draw_plot(pa05, x = -0.02, y = 0.8, width = 0.1, height = 0.18) +
  draw_plot(pa06, x = -0.02, y = 0, width = 0.1, height = 0.18) +
  draw_plot(pa07, x = -0.02, y = 0.2, width = 0.1, height = 0.18) +
  #draw_plot(pa08, x = 0.2, y = 0, width = 0.1, height = 0.18) +
  draw_plot(pa09, x = 0.13, y = 0.6, width = 0.1, height = 0.18) +
  draw_plot(pa10, x = 0.13, y = 0.2, width = 0.1, height = 0.18) +
  draw_plot(pa11, x = 0.13, y = 0.4, width = 0.1, height = 0.18) +
  draw_plot(pa12, x = 0.13, y = 0, width = 0.1, height = 0.18) +
  draw_plot(pa13, x = 0.13, y = 0.8, width = 0.1, height = 0.18) +
  draw_plot(f2a, x = 0.3, y = 0.47, width = 0.55, height = 0.53) +
  #draw_plot(legend, x = 0.15, y = 0.34, width = 0.1, height = 0.2) +
  draw_plot(f2b, x = 0.85, y = 0.47, width = 0.15, height = 0.53) +
  draw_plot(f2c, x = 0.3, y = 0, width = 0.55, height = 0.47) +
  draw_plot(f2d, x = 0.85, y = 0, width = 0.15, height = 0.47)

```{r}
#| eval: false

save_plot(filename = "../manuscript/Coauthors/PNAS/Revision/Figure_2_Covariance_matrices_and_posterior_weights.svg", plot = last_plot(), base_height = 6, base_width = 8.5)
```

@@

Change these to look at S_AP, S_DS, S_2_no; S_1_row and S_1_col separately

@@

get_gxe_summary <- function(gxe_list, n_cond = 8, n_sep = 2) {
  AP <- one_gxe_summary(gxe_list$S_AP, n_cond = n_cond, n_sep = n_sep) %>%
    rename(Sign = n)
  DS <- one_gxe_summary(gxe_list$S_DS, n_cond = n_cond, n_sep = n_sep) %>%
    rename(Magnitude = n)
  ND <- one_gxe_summary(gxe_list$S_2_no, n_cond = n_cond, n_sep = n_sep) %>%
    rename(None = n)
  gxe_n <- AP %>%
    full_join(DS, by = "Region_comparison") %>%
    full_join(ND, by = "Region_comparison") %>%
    pivot_longer(cols = -Region_comparison, names_to = "Type_of_GxE", 
                 values_to = "n") %>%
    pivot_wider(names_from = "Region_comparison", values_from = "n")
  gxe_frac <- gxe_n %>%
    mutate(North_North = round(North_North/sum(North_North), digits = 3),
           North_Texas = round(North_Texas/sum(North_Texas), digits = 3),
           Texas_Texas = round(Texas_Texas/sum(Texas_Texas), digits = 3))
  return(list(gxe_counts = gxe_n, gxe_fractions = gxe_frac))
}

@@

```{r}
get_gxe_summary(gxe_gg) 
get_gxe_summary(gxe_mg, n_cond = 7) 
get_gxe_summary(gxe_gamg, n_sep = 4) 
print("flowering")
get_gxe_summary(gxe_gf) 
get_gxe_summary(gxe_mf, n_cond = 6) 
get_gxe_summary(gxe_gamf, n_sep = 4) 
```

@@

```{r}
 pvdiv_plot_cor(corrmatrix = gxe_gamf$S_AP, legend = TRUE) + 
  labs(subtitle = " ")
```

@@

```{r}
#| eval: false

@@

Comparison with QTL for the four-way mapping cross for green up (GR50), flowering (FL50), three functions of weather variables for green-up (from Figure 2; Both: Temp Ave, 1d; Both: Daylength change (s) 1d prior; Gulf: Daylength (14d prior), and two functions of weather variables for flowering (from Figure 2: Midwest: day length change (s) 2d prior; Gulf: cumulative rainfall, 7d).

Li ran the QTL analysis 2024-02-14. Here Alice combines the QTL with the mash SNP effects to build the analysis for Figure 4.

It's an open question how to combine these into an informative main text figure.

The full QTL Lod plots for flowering & greenup phenotypes should go in the SI. Here, try plotting just the significant chromosomes, with a QTL above the permutation threshold.

#### Figure 4A

LOD QTL

Adapted from Analysis_v0.7_QTL_interval_lod_plot.Rmd

Make sure to Load mash objects at start of Step 5

```{r}
library(qtl)
library(qtl2)
# library(plyr)
# library(tidyverse)
library(qtl2convert)
#library(cowplot)
```

```{r}
workingdir <- "~/Github/pvdiv-phenology-gxe/phenology-fourway/5-QTLmappingForAlice/"
load(file.path(workingdir, 'cross_reduced_marker_FL50_phenotypes_Alice_7_SITES.RData'))
load(file.path(workingdir, 'scan1Results_FL50_phenotypes_Alice_7_SITES.RData'))

map  =  lapply(pull.map(cross), function(x) x[1,])
###get the threshold value for each model and output the QTL intervals
threshold_full = summary(s1permfull,alpha = 0.05)
threshold_reduced = summary(s1permreduced,alpha = 0.05)
```

```{r}
###  g x e
gxe = s1full - s1reduced
gxe_perm = threshold_full - threshold_reduced ###or gxe_perm = summary(gxe, alpha=0.05)

###get the peak for full model and gxe model
peaks_full = find_peaks(s1full, map, threshold = threshold_full, drop = 1.5, peakdrop = 5)
peaks_gxe = find_peaks(gxe, map, threshold = gxe_perm, drop = 1.5)

flank_marker_full = peaks_full %>% mutate(flank_lo = find_marker(map, chr, ci_lo), flank_hi=find_marker(map, chr, ci_hi))
# write.csv(flank_marker_full,'QTLsWithFlankMarkers_FL50_phenotypes_Alice_7_SITES_Full_model.csv') ###change the name you want

flank_marker_gxe = peaks_gxe %>% mutate(flank_lo = find_marker(map, chr, ci_lo), flank_hi=find_marker(map, chr, ci_hi))
# write.csv(flank_marker_gxe,'QTLsWithFlankMarkers_FL50_phenotypes_Alice_7_SITES_gxe_model.csv')

###########
```

```{r}
get_post_weights <- function(m){
  log10bf_df <- switchgrassGWAS:::get_log10bf(m = m) %>%
      as.data.frame() %>%
      rownames_to_column(var = "value") %>%
      dplyr::mutate(value = as.integer(.data$value)) %>%
      as_tibble() %>%
      left_join(switchgrassGWAS:::get_marker_df(m = m), by = "value") %>%
      dplyr::rename(log10BayesFactor = .data$V1) %>%
      dplyr::select(-.data$value)

  pw_df <- as_tibble(m$posterior_weights)
  pw_df <- switchgrassGWAS:::get_marker_df(m = m) %>%
    bind_cols(pw_df)

  ggman_df <- pw_df %>%
    #left_join(m_meta) %>%
    tidyr::separate(.data$Marker, into = c("Chr", "Pos"), remove = FALSE, sep = "_",
             extra = "merge", convert = TRUE) %>%
    left_join(log10bf_df, by = "Marker") %>%
    dplyr::arrange(.data$Chr, .data$Pos) 
  return(ggman_df)
}

pw_gg <- get_post_weights(m_gulf_gr_hyp)
pw_gf <- get_post_weights(m_gulf_fl_hyp) %>%
  mutate(Subpop = "Gulf")

pw_mg <- get_post_weights(m_midw_gr_hyp)
pw_mf <- get_post_weights(m_midw_fl_hyp) %>%
  mutate(Subpop = "Midwest")

pw_gamg <- get_post_weights(m_gam_gr_hyp)
pw_gamf <- get_post_weights(m_gam_fl_hyp) %>%
  mutate(Subpop = "Both")
```

Decide on cutoff to use for rug plots and enrichment analysis: SNPs sig in 1% tail or BF>2, whichever is stricter

1% tail is stricter in all cases except Midwest flowering, where BF>2 is stricter, equivalent to a 0.49% tail (0.995119 quantile).

What is the Bayes Factor at the 99th quantile? If it's not above 2, what is the quantile where the BF > 2?
```{r}
quantile(pw_gf$log10BayesFactor, 0.99)  # Gulf: Use 2.56 BF
quantile(pw_mf$log10BayesFactor, 0.99)  # Midwest: Use 2 BF
quantile(pw_gamf$log10BayesFactor, 0.99)  # GAM: Use 2.65 BF
quantile(pw_mf$log10BayesFactor, 0.995119)

quantile(pw_gg$log10BayesFactor, 0.99)  # Gulf: Use 2.0989 BF
quantile(pw_mg$log10BayesFactor, 0.99)  # Midwest: Use 4.6831 BF
quantile(pw_gamg$log10BayesFactor, 0.99)  # GAM: Use 3.0865 BF
```

```{r}
rug_gf <- pw_gf %>%
  filter(log10BayesFactor > 2.559618) %>% 
  mutate(Pos = Pos/1000000, 
         LOD = -0.5)
rug_gamf <- pw_gamf %>%
  filter(log10BayesFactor > 2.64938) %>% 
  mutate(Pos = Pos/1000000, 
         LOD = -0.5)
rug_mf <- pw_mf %>%
  filter(log10BayesFactor > 2) %>% 
  mutate(Pos = Pos/1000000, 
         LOD = -0.5)

rug_gg <- pw_gg %>%
  filter(log10BayesFactor > 2.0989) %>% 
  mutate(Pos = Pos/1000000, 
         LOD = -0.5)
rug_gamg <- pw_gamg %>%
  filter(log10BayesFactor > 3.0865) %>% 
  mutate(Pos = Pos/1000000, 
         LOD = -0.5)
rug_mg <- pw_mg %>%
  filter(log10BayesFactor > 4.6831) %>% 
  mutate(Pos = Pos/1000000, 
         LOD = -0.5)
```

```{r}
threshold_df <- threshold_full %>% 
  as_tibble(rownames = "Threshold") %>%
  pivot_longer(GR50:crain_7d, names_to = "Phenotype", values_to = "LOD") %>%
  mutate(Phe = case_when(Phenotype %in% c("GR50", "tave_1d", "dyln_14d_prior", "dyln_change_sec") ~ "greenup",
                         Phenotype %in% c("FL50", "dyln_change_sec_2d_prior", "crain_7d") ~ "flowering", 
                         TRUE ~ NA_character_))
